---
title: "Egg-clutch biomechanics affect escape-hatching behavior: A comparison of congeneric treefrogs "
author: "Brandon Güell"
date: "3/15/2023"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
---

# Introduction

> The data processing and statistical analyses shown here are from the Guell et al. "Egg-clutch biomechanics affect escape-hatching behavior and performance". Analyses are shows in chronological order as reported in the manuscript. We show details on importing, manipulating, summarizing, analyzing (testing hypotheses), and visualizing the data. See manuscript for details on background, research questions, hypothesis, methods, etc..


PDF Knit settings
```{r global_options}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

# Clutch Biomechanics

Import Datasheets

Manual measurements
```{r}
library(readxl)
library(dplyr)
library(ggplot2)
library(scales)

biomechmeasurements <- read_excel("~/Desktop/Publications/Manuscripts/In Prep/Comparative Egg Clutch Biomechanics/Comparative Egg Clutch Biomechanics.xlsx",
    sheet = "Manual Vibration Measurements", na = "NA")
biomechmeasurements

#subsets of data
ringdown_only_biomechmeasurements = biomechmeasurements %>%
  filter(Vibration_Signal_Portion == "Ringdown")
ringdown_only_biomechmeasurements

entire_disturbance_biomechmeasurements = biomechmeasurements %>%
  filter(Vibration_Signal_Portion == "Entire disturbance")
entire_disturbance_biomechmeasurements
```

## Egg clutch Morphology
Table 1 summary
```{r}
clutch_size_sum = biomechmeasurements %>%
  filter(Excitation_Test=="Pendulum impact" & Vibration_Signal_Portion =="Ringdown" & Impact_Distance_cm==1) %>%
  group_by(Species) %>%
summarise(mean = mean(Clutch_Size, na.rm=T), SE= sd(Clutch_Size, na.rm=T)/sqrt(length(Clutch_Size)), SD = sd(Clutch_Size, na.rm=T), min = min(Clutch_Size, na.rm=T), max = max(Clutch_Size, na.rm=T), sample_size_unique=n_distinct(Clutch_Size, na.rm=T), sample_size=n())
clutch_size_sum

clutch_length_sum = biomechmeasurements %>%
  filter(Excitation_Test=="Pendulum impact" & Vibration_Signal_Portion =="Ringdown" & Impact_Distance_cm==1) %>%
  group_by(Species) %>%
summarise(mean = mean(Clutch_Length_mm, na.rm=T), SE= sd(Clutch_Length_mm, na.rm=T)/sqrt(length(Clutch_Length_mm)), SD = sd(Clutch_Length_mm, na.rm=T), min = min(Clutch_Length_mm, na.rm=T), max = max(Clutch_Length_mm, na.rm=T), sample_size_unique=n_distinct(Clutch_Length_mm, na.rm=T), sample_size=n())
clutch_length_sum

clutch_width_sum = biomechmeasurements %>%
  filter(Excitation_Test=="Pendulum impact" & Vibration_Signal_Portion =="Ringdown" & Impact_Distance_cm==1) %>%
  group_by(Species) %>%
summarise(mean = mean(Clutch_Width_mm, na.rm=T), SE= sd(Clutch_Width_mm, na.rm=T)/sqrt(length(Clutch_Width_mm)), SD = sd(Clutch_Width_mm, na.rm=T), min = min(Clutch_Width_mm, na.rm=T), max = max(Clutch_Width_mm, na.rm=T), sample_size_unique=n_distinct(Clutch_Width_mm, na.rm=T), sample_size=n())
clutch_width_sum

clutch_thickness_sum = biomechmeasurements %>%
  filter(Excitation_Test=="Pendulum impact" & Vibration_Signal_Portion =="Ringdown" & Impact_Distance_cm==1) %>%
  group_by(Species) %>%
summarise(mean = mean(Clutch_Thickness_mm, na.rm=T), SE= sd(Clutch_Thickness_mm, na.rm=T)/sqrt(length(Clutch_Thickness_mm)), SD = sd(Clutch_Thickness_mm, na.rm=T), min = min(Clutch_Thickness_mm, na.rm=T), max = max(Clutch_Thickness_mm, na.rm=T), sample_size_unique=n_distinct(Clutch_Thickness_mm, na.rm=T), sample_size=n())
clutch_thickness_sum
```

### STATS

Clutch Size
```{r}
clutchdata= biomechmeasurements %>%
  group_by(Clutch, Species) %>%
  summarise(mean = mean(Clutch_Size, na.rm=T), SE= sd(Clutch_Size, na.rm=T)/sqrt(length(Clutch_Size)), SD = sd(Clutch_Size, na.rm=T), max = max(Clutch_Size, na.rm=T), min = min(Clutch_Size, na.rm=T), sample_size_unique=n_distinct(Clutch_Size, na.rm=T), sample_size=n())
clutchdata

m=lm(mean~Species, data=clutchdata)
hist(clutchdata$mean)

plot(fitted(m), residuals(m))
hist(residuals(m))
qqnorm(residuals(m))
qqline(residuals(m))
shapiro.test(residuals(m))

t.test(mean~Species, data=clutchdata)
```

Welch Two Sample t-test

data:  mean by Species
*t = 2.2154, df = 30.392, p-value = 0.03436*
alternative hypothesis: true difference in means between group A. callidryas and group A. spurrelli is not equal to 0
95 percent confidence interval:
  0.5401357 13.1951584
sample estimates:
mean in group A. callidryas  mean in group A. spurrelli 
                   40.11765                    33.25000 


Clutch length
```{r}
clutchdata2= biomechmeasurements %>%
  group_by(Clutch, Species) %>%
  summarise(mean = mean(Clutch_Length_mm, na.rm=T), SE= sd(Clutch_Length_mm, na.rm=T)/sqrt(length(Clutch_Length_mm)), SD = sd(Clutch_Length_mm, na.rm=T), max = max(Clutch_Length_mm, na.rm=T), min = min(Clutch_Length_mm, na.rm=T), sample_size_unique=n_distinct(Clutch_Length_mm, na.rm=T), sample_size=n())
clutchdata2

m=lm(mean~Species, data=clutchdata2)
hist(clutchdata2$mean)

plot(fitted(m), residuals(m))
hist(residuals(m))
qqnorm(residuals(m))
qqline(residuals(m))
shapiro.test(residuals(m))

t.test(mean~Species, data=clutchdata2)
```

Welch Two Sample t-test

data:  mean by Species
*t = 1.0203, df = 29.994, p-value = 0.3157*
alternative hypothesis: true difference in means between group A. callidryas and group A. spurrelli is not equal to 0
95 percent confidence interval:
 -2.593829  7.773241
sample estimates:
mean in group A. callidryas  mean in group A. spurrelli 
                   46.26471                    43.67500 


Clutch width
```{r}
clutchdata3= biomechmeasurements %>%
  group_by(Clutch, Species) %>%
  summarise(mean = mean(Clutch_Width_mm, na.rm=T), SE= sd(Clutch_Width_mm, na.rm=T)/sqrt(length(Clutch_Width_mm)), SD = sd(Clutch_Width_mm, na.rm=T), max = max(Clutch_Width_mm, na.rm=T), min = min(Clutch_Width_mm, na.rm=T), sample_size_unique=n_distinct(Clutch_Width_mm, na.rm=T), sample_size=n())
clutchdata3

m=lm(mean~Species, data=clutchdata3)
hist(clutchdata3$mean)

plot(fitted(m), residuals(m))
hist(residuals(m))
qqnorm(residuals(m))
qqline(residuals(m))
shapiro.test(residuals(m))

t.test(mean~Species, data=clutchdata3)
```

Welch Two Sample t-test

data:  mean by Species
*t = 2.6378, df = 30.223, p-value = 0.01306*
alternative hypothesis: true difference in means between group A. callidryas and group A. spurrelli is not equal to 0
95 percent confidence interval:
 0.798982 6.271606
sample estimates:
mean in group A. callidryas  mean in group A. spurrelli 
                   23.93529                    20.40000 


Clutch thickness
```{r}
clutchdata4= biomechmeasurements %>%
  group_by(Clutch, Species) %>%
  summarise(mean = mean(Clutch_Thickness_mm, na.rm=T), SE= sd(Clutch_Thickness_mm, na.rm=T)/sqrt(length(Clutch_Thickness_mm)), SD = sd(Clutch_Thickness_mm, na.rm=T), max = max(Clutch_Thickness_mm, na.rm=T), min = min(Clutch_Thickness_mm, na.rm=T), sample_size_unique=n_distinct(Clutch_Thickness_mm, na.rm=T), sample_size=n())
clutchdata4

m=lm(mean~Species, data=clutchdata4)
hist(clutchdata4$mean)

plot(fitted(m), residuals(m))
hist(residuals(m))
qqnorm(residuals(m))
qqline(residuals(m))
shapiro.test(residuals(m))

t.test(mean~Species, data=clutchdata4)
```

Welch Two Sample t-test

data:  mean by Species
*t = 3.9847, df = 30.9, p-value = 0.0003828*
alternative hypothesis: true difference in means between group A. callidryas and group A. spurrelli is not equal to 0
95 percent confidence interval:
 0.8046443 2.4924145
sample estimates:
mean in group A. callidryas  mean in group A. spurrelli 
                   9.423529                    7.775000 


# MATLAB Model Fit
```{r}
library(readxl)
library(dplyr)
biomech <- read_excel("~/Desktop/Publications/Manuscripts/In Prep/Comparative Egg Clutch Biomechanics/Comparative Egg Clutch Biomechanics.xlsx", 
    sheet = "Biomechanics Code Output")

biomech

# remove observations based on criteria - calculate outliers at 3-s.d. from mean
biomech = biomech %>%
  filter(NRMSE < 0.5) %>%
  filter(Time_Constant_s < 3*sd(biomech$Time_Constant_s)) %>%
  filter(Frequency_Hz < 3*sd(biomech$Frequency_Hz)) %>%
  filter(Frequency_Hz > 0) %>%
  filter(Time_Constant_s > 0) %>%
  filter(Period_s > 0) %>%
  filter(Time_Constant_s < 0.5) 
biomech

biomech_as_only = biomech %>%
  filter(Species =="A. spurrelli")

biomech_ac_only = biomech %>%
  filter(Species =="A. callidryas")
```

## Model Fit Accuracy
Table 2 Summary Statistics
```{r}
library(dplyr)
biomech_summary3 = biomech %>%
  group_by(Species, Excitation_Test) %>%
summarise(mean = mean(Time_Between_Samples_s, na.rm=T), SE= sd(Time_Between_Samples_s, na.rm=T)/sqrt(length(Time_Between_Samples_s)), SD = sd(Time_Between_Samples_s, na.rm=T), max = max(Time_Between_Samples_s, na.rm=T), min = min(Time_Between_Samples_s, na.rm=T), sample_size_unique=n_distinct(Time_Between_Samples_s, na.rm=T), sample_size=n())
biomech_summary3

library(dplyr)
biomech_summary6 = biomech %>%
  group_by(Species, Excitation_Test) %>%
summarise(mean = mean(NRMSE, na.rm=T), SE= sd(NRMSE, na.rm=T)/sqrt(length(NRMSE)), SD = sd(NRMSE, na.rm=T), max = max(NRMSE, na.rm=T), min = min(NRMSE, na.rm=T), sample_size_unique=n_distinct(NRMSE, na.rm=T), sample_size=n())
biomech_summary6
```

Visualizations
```{r}
library(ggplot2)
library(scales)
library(lemon)

biomechplot3= ggplot(biomech, aes(x=factor(Excitation_Test, level = c("Pendulum impact 1", "Pendulum impact 2", "Small water drop", "Large water drop")), y=Time_Between_Samples_s)) + #add the axis
  geom_boxplot(aes(color=Species), outlier.shape = NA, fill = NA) +
  geom_point(aes(fill=Species), position = position_jitterdodge(.2), size = 1, color = "black", pch=21) +
  labs(y="Window length (s)", x=element_blank()) + #creat labels... 
  theme_classic(base_size=8) +
  theme(legend.text = element_text(size = 8), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 8), axis.ticks = element_line(color = "black"), axis.text.x = element_blank()) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= "none") +
  scale_y_continuous(breaks= pretty_breaks(n=6)) + # allows for making "pretty" breaks on y axis... shows 5 breaks
  scale_color_manual(values = c("red3", "darkblue")) +
  scale_fill_manual(values = c("red3", "darkblue")) +
  facet_grid(cols = vars(Excitation_Test_Type), as.table = T, scales = "free", space = "free")
biomechplot3

biomechplot6= ggplot(biomech, aes(x=factor(Excitation_Test, level = c("Pendulum impact 1", "Pendulum impact 2", "Small water drop", "Large water drop")), y=NRMSE)) + #add the axis
  geom_boxplot(aes(color=Species), outlier.shape = NA, fill = NA) +
  geom_point(aes(fill=Species), position = position_jitterdodge(.2), size = 1, color = "black", pch=21) +
  labs(y="NRMSE", x=element_blank()) + #creat labels... 
  theme_classic(base_size=8) +
  theme(legend.text = element_text(size = 8), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 8), axis.ticks = element_line(color = "black")) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= "none", strip.background = element_blank(), strip.text = element_blank()) +
  scale_y_continuous(breaks= pretty_breaks(n=6)) + # allows for making "pretty" breaks on y axis... shows 5 breaks
  scale_color_manual(values = c("red3", "darkblue")) +
  scale_fill_manual(values = c("red3", "darkblue")) +
  facet_grid(cols = vars(Excitation_Test_Type), as.table = T, scales = "free", space = "free")
biomechplot6
```

Composite Figure S1
```{r}
library(gridExtra)
library(cowplot)

#  @ portrait at 7.08661 x 8.26772 max for JEB
# 3.46x6 portrait looks good

SuppFigure1 = plot_grid(biomechplot3, biomechplot6, labels = "AUTO", label_x = -.01, label_y = 1, scale =1, ncol=1, nrow = 2, label_size = 12, align = "v")
SuppFigure1
```


### STATS

#### Window Length
```{r}
library(lme4)
library(multcomp)

sup1_fullmodel = lmer(Time_Between_Samples_s~Excitation_Test*Species +  (1|Clutch), data=biomech)
sup1_nointeraction = lmer(Time_Between_Samples_s~Excitation_Test+Species + (1|Clutch), data=biomech)
sup1_noexcitationtest = lmer(Time_Between_Samples_s~Species + (1|Clutch), data=biomech)
sup1_nospecies = lmer(Time_Between_Samples_s~Excitation_Test + (1|Clutch), data=biomech)


#AIC model comparison
library(AICcmodavg)
print(aictab(list(sup1_fullmodel,  sup1_nointeraction, sup1_noexcitationtest, sup1_nospecies), c("sup1_fullmodel", "sup1_nointeraction", "sup1_noexcitationtest", "sup1_nospecies")), LL = FALSE)

summary(sup1_distnace)

anova(sup1_fullmodel, sup1_nointeraction, test = "Chisq")
anova(sup1_fullmodel, sup1_nospecies, test = "Chisq")
anova(sup1_fullmodel, sup1_noexcitationtest, test = "Chisq")

library(multcomp)
summary(glht(sup1_fullmodel, linfct=mcp(Excitation_Test="Tukey")))
```
EFFECT OF INTERACTION:
* **X = 13.982; p-value = 0.002929 **

EFFECT OF SPECIES:
* **X = 294.78; p-value  < 2.2e-16 **

EFFECT OF EXCITATION TEST:
* **X = 20.567; p-value = 0.002194**

	 Simultaneous Tests for General Linear Hypotheses

Multiple Comparisons of Means: Tukey Contrasts


Fit: lmer(formula = Time_Between_Samples_s ~ Excitation_Test * Species + 
    (1 | Clutch), data = biomech)

Linear Hypotheses:
                                            Estimate Std. Error z value Pr(>|z|)    
Pendulum impact 1 - Large water drop == 0  -0.019926   0.009743  -2.045   0.1714    
Pendulum impact 2 - Large water drop == 0   0.022937   0.010277   2.232   0.1146    
Small water drop - Large water drop == 0    0.006021   0.009913   0.607   0.9297    
Pendulum impact 2 - Pendulum impact 1 == 0  0.042863   0.010401   4.121   <0.001 ***
Small water drop - Pendulum impact 1 == 0   0.025947   0.009992   2.597   0.0466 *  
Small water drop - Pendulum impact 2 == 0  -0.016916   0.010536  -1.606   0.3752   



Within Species
```{r}
library(lme4)
library(multcomp)
#As
sup1_fullmodel = lmer(Time_Between_Samples_s~Excitation_Test +  (1|Clutch), data=biomech_as_only)
sup1_noexcitationtest = lmer(Time_Between_Samples_s~ + (1|Clutch), data=biomech_as_only)

anova(sup1_fullmodel, sup1_noexcitationtest, test = "Chisq")
summary(glht(sup1_fullmodel, linfct=mcp(Excitation_Test="Tukey")))


#Ac
sup1_fullmodel = lmer(Time_Between_Samples_s~Excitation_Test +  (1|Clutch), data=biomech_ac_only)
sup1_noexcitationtest = lmer(Time_Between_Samples_s~ + (1|Clutch), data=biomech_ac_only)

anova(sup1_fullmodel, sup1_noexcitationtest, test = "Chisq")

library(multcomp)
summary(glht(sup1_fullmodel, linfct=mcp(Excitation_Test="Tukey")))
```

A. spurrelli:
EFFECT OF EXCITATION TYPE:
* **X = 48.781; p-value  1.453e-10 **

Pendulum impact 1 - Large water drop == 0   0.015811   0.002771   5.705  < 0.001 ***
Pendulum impact 2 - Large water drop == 0   0.006017   0.002790   2.157  0.13547    
Small water drop - Large water drop == 0   -0.002680   0.002822  -0.950  0.77785    
Pendulum impact 2 - Pendulum impact 1 == 0 -0.009794   0.002677  -3.659  0.00137 ** 
Small water drop - Pendulum impact 1 == 0  -0.018491   0.002741  -6.746  < 0.001 ***
Small water drop - Pendulum impact 2 == 0  -0.008697   0.002760  -3.151  0.00884 **

A. callidryas:
EFFECT OF EXCITATION TYPE:
* **X = 18.267; p-value  = 0.0003875 **

Pendulum impact 1 - Large water drop == 0  -0.022404   0.010640  -2.106   0.1511    
Pendulum impact 2 - Large water drop == 0   0.026850   0.011276   2.381   0.0805 .  
Small water drop - Large water drop == 0    0.002038   0.010822   0.188   0.9976    
Pendulum impact 2 - Pendulum impact 1 == 0  0.049254   0.011382   4.327   <0.001 ***
Small water drop - Pendulum impact 1 == 0   0.024442   0.010920   2.238   0.1129    
Small water drop - Pendulum impact 2 == 0  -0.024812   0.011553  -2.148   0.1382 



#### NRMSE
```{r}
sup1b_fullmodel = lmer(NRMSE~Excitation_Test*Species +  (1|Clutch), data=biomech)
sup1b_nointeraction = lmer(NRMSE~Excitation_Test+Species + (1|Clutch), data=biomech)
sup1b_noexcitationtest = lmer(NRMSE~Species + (1|Clutch), data=biomech)
sup1b_nospecies = lmer(NRMSE~Excitation_Test + (1|Clutch), data=biomech)


#AIC model comparison
library(AICcmodavg)
print(aictab(list(sup1b_fullmodel,  sup1b_nointeraction, sup1b_noexcitationtest, sup1b_nospecies), c("sup1b_fullmodel", "sup1b_nointeraction", "sup1b_noexcitationtest", "sup1b_nospecies")), LL = FALSE)

anova(sup1b_fullmodel, sup1b_nointeraction, test = "Chisq")
anova(sup1b_fullmodel, sup1b_nospecies, test = "Chisq")
anova(sup1b_fullmodel, sup1b_noexcitationtest, test = "Chisq")

library(multcomp)
summary(glht(sup1b_fullmodel, linfct=mcp(Excitation_Test="Tukey")))
```

EFFECT OF INTERACTION:
* **X = 8.0744; p-value = 0.0445 **

EFFECT OF SPECIES:
* **X = 39.079; p-value 6.71e-08 **

EFFECT OF EXCITATION TEST:
* **X = 65.76; p-value = 3.017e-12 **



	 Simultaneous Tests for General Linear Hypotheses

Multiple Comparisons of Means: Tukey Contrasts


Fit: lmer(formula = NRMSE ~ Excitation_Test * Species + (1 | Clutch), 
    data = biomech)

Linear Hypotheses:
                                            Estimate Std. Error z value Pr(>|z|)    
Pendulum impact 1 - Large water drop == 0  -0.059359   0.012275  -4.836   <0.001 ***
Pendulum impact 2 - Large water drop == 0   0.007867   0.012941   0.608   0.9295    
Small water drop - Large water drop == 0    0.032115   0.012529   2.563   0.0508 .  
Pendulum impact 2 - Pendulum impact 1 == 0  0.067226   0.013036   5.157   <0.001 ***
Small water drop - Pendulum impact 1 == 0   0.091475   0.012601   7.260   <0.001 ***
Small water drop - Pendulum impact 2 == 0   0.024249   0.013263   1.828   0.2599 








## Clutch Morhpometrics 
```{r}
clutch_size_and_thickness= ggplot(biomech, aes(x=Clutch_Size, y=Clutch_Thickness_mm)) + #add the axis
  geom_smooth(aes(color=Species, group=interaction(Species,Excitation_Test), linetype=Species), method = "glm", se=F, size = 1 ) +
  geom_point(aes(fill=Species), position = position_jitterdodge(.2), size = 1, color="black", pch=21) +
  labs(y="Clutch thickness", x="Clutch size") + #creat labels... 
  theme_classic(base_size=8) +
  theme(legend.text = element_text(size = 8), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 8), axis.ticks = element_line(color = "black")) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= c(.025,.95)) +
  scale_y_continuous(breaks= pretty_breaks(n=6)) + # allows for making "pretty" breaks on y axis... shows 5 breaks
  scale_color_manual(values = c("red3", "darkblue")) +
  scale_fill_manual(values = c("red3", "darkblue")) 
clutch_size_and_thickness

correlationstuff = glm(Clutch_Size~Clutch_Thickness_mm, data=biomech, family=Gamma(link=log))
correlationstuff
summary(correlationstuff)
Anova(correlationstuff)

#calculate R-squared like this
with(summary(correlationstuff), 1 - deviance/null.deviance)
```

R^2 = 0.4110576





## Modal Analyses

### Frequency

Summarize data
```{r}
library(dplyr)
biomech_summary1 = biomech %>%
  group_by(Species, Excitation_Test, Excitation_Test_Type, Clutch) %>%
summarise(mean = mean(Frequency_Hz, na.rm=T), SE= sd(Frequency_Hz, na.rm=T)/sqrt(length(Frequency_Hz)), SD = sd(Frequency_Hz, na.rm=T), max = max(Frequency_Hz, na.rm=T), min = min(Frequency_Hz, na.rm=T), sample_size_unique=n_distinct(Frequency_Hz, na.rm=T), sample_size=n())
biomech_summary1

biomech_summary1.5 = biomech %>%
  group_by(Species, Excitation_Test_Type, Clutch) %>%
summarise(mean = mean(Frequency_Hz, na.rm=T), SE= sd(Frequency_Hz, na.rm=T)/sqrt(length(Frequency_Hz)), SD = sd(Frequency_Hz, na.rm=T), max = max(Frequency_Hz, na.rm=T), min = min(Frequency_Hz, na.rm=T), sample_size_unique=n_distinct(Frequency_Hz, na.rm=T), sample_size=n())
biomech_summary1.5

biomech_summary2 = biomech %>%
  group_by(Species, Excitation_Test_Type) %>%
summarise(mean = mean(Frequency_Hz, na.rm=T), SE= sd(Frequency_Hz, na.rm=T)/sqrt(length(Frequency_Hz)), SD = sd(Frequency_Hz, na.rm=T), max = max(Frequency_Hz, na.rm=T), min = min(Frequency_Hz, na.rm=T), sample_size_unique=n_distinct(Frequency_Hz, na.rm=T), sample_size=n())
biomech_summary2

biomech_summary2.5 = biomech %>%
  group_by(Species, Excitation_Test) %>%
summarise(mean = mean(Frequency_Hz, na.rm=T), SE= sd(Frequency_Hz, na.rm=T)/sqrt(length(Frequency_Hz)), SD = sd(Frequency_Hz, na.rm=T), max = max(Frequency_Hz, na.rm=T), min = min(Frequency_Hz, na.rm=T), sample_size_unique=n_distinct(Frequency_Hz, na.rm=T), sample_size=n())
biomech_summary2.5

biomech_summary3 = biomech %>%
  group_by(Species, Clutch) %>%
summarise(mean = mean(Frequency_Hz, na.rm=T), SE= sd(Frequency_Hz, na.rm=T)/sqrt(length(Frequency_Hz)), SD = sd(Frequency_Hz, na.rm=T), max = max(Frequency_Hz, na.rm=T), min = min(Frequency_Hz, na.rm=T), sample_size_unique=n_distinct(Frequency_Hz, na.rm=T), sample_size=n())
biomech_summary3

biomech_summary4 = biomech %>%
  group_by(Species) %>%
summarise(mean = mean(Frequency_Hz, na.rm=T), SE= sd(Frequency_Hz, na.rm=T)/sqrt(length(Frequency_Hz)), SD = sd(Frequency_Hz, na.rm=T), max = max(Frequency_Hz, na.rm=T), min = min(Frequency_Hz, na.rm=T), sample_size_unique=n_distinct(Frequency_Hz, na.rm=T), sample_size=n())
biomech_summary4
```

> The significant species × excitation type interaction was due to the lower frequencies excited by pendulum impacts in A. spurrelli which were on average ~30 Hz lower than those by water drops 

Clutch Morphometrics
```{r}
frequency_vs_clutchsize_raw= ggplot(biomech, aes(x=Clutch_Size, y=Frequency_Hz)) + #add the axis
  geom_smooth(aes(color=Species, group=interaction(Species,Excitation_Test), linetype=Species), method = "glm", se=F, size = 1 ) +
  geom_point(aes(fill=Species), position = position_jitterdodge(.2), size = 1, color="black", pch=21) +
  labs(y="Frequency (Hz)", x="Clutch size") + #creat labels... 
  theme_classic(base_size=8) +
  theme(legend.text = element_text(size = 8), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 8), axis.ticks = element_line(color = "black")) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= c(.025,.95)) +
  scale_y_continuous(breaks= pretty_breaks(n=6)) + # allows for making "pretty" breaks on y axis... shows 5 breaks
  scale_color_manual(values = c("red3", "darkblue")) +
  scale_fill_manual(values = c("red3", "darkblue")) 
frequency_vs_clutchsize_raw

frequency_vs_clutchsize_raw= ggplot(biomech, aes(x=Clutch_Thickness_mm, y=Frequency_Hz)) + #add the axis
  geom_smooth(aes(color=Species, group=interaction(Species,Excitation_Test), linetype=Species), method = "glm", se=F, size = 1 ) +
  geom_point(aes(fill=Species), position = position_jitterdodge(.2), size = 1, color="black", pch=21) +
  labs(y="Frequency (Hz)", x="Clutch Thickness") + #creat labels... 
  theme_classic(base_size=8) +
  theme(legend.text = element_text(size = 8), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 8), axis.ticks = element_line(color = "black")) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= c(.025,.95)) +
  scale_y_continuous(breaks= pretty_breaks(n=6)) + # allows for making "pretty" breaks on y axis... shows 5 breaks
  scale_color_manual(values = c("red3", "darkblue")) +
  scale_fill_manual(values = c("red3", "darkblue")) 
frequency_vs_clutchsize_raw

library(car)
# effect of clutch size?
m2 = lmer(Frequency_Hz~Species*Clutch_Size*Excitation_Test +  (1|Clutch), data=biomech)
Anova(m2)

m2 = lmer(Frequency_Hz~Species*Clutch_Thickness_mm +  (1|Clutch), data=biomech)
Anova(m2)
```


#### Visualization
```{r}
library(ggplot2)
library(scales)
frequency_raw_perclutch= ggplot(biomech_summary1, aes(x=factor(Excitation_Test, level = c("Pendulum impact 1", "Pendulum impact 2", "Small water drop", "Large water drop")), y=mean)) + #add the axis
  geom_boxplot(aes(color=Species), outlier.shape = NA, fill = NA) +
  geom_point(aes(fill=Species), position = position_jitterdodge(.2), size = 1, color="black", pch=21) +
  labs(y="Frequency (Hz)", x=element_blank()) + #creat labels... 
  theme_classic(base_size=8) +
  theme(legend.text = element_text(size = 8, face = "italic"), legend.title = element_blank(),  legend.background = element_blank(), axis.text = element_text(color="black", size = 8), axis.ticks = element_line(color = "black"), strip.text = element_text(size = 8)) +
  theme(legend.title.align=0.5, axis.text.x = element_blank(), legend.position = c(.85, .95)) +
  scale_y_continuous(breaks= pretty_breaks(n=7)) +
  scale_color_manual(values = c("red3", "darkblue")) +
  scale_fill_manual(values = c("red3", "darkblue")) +
  facet_grid(cols = vars(Excitation_Test_Type), as.table = T, scales = "free", space = "free")
frequency_raw_perclutch

frequency_raw= ggplot(biomech, aes(x=factor(Excitation_Test, level = c("Pendulum impact 1", "Pendulum impact 2", "Small water drop", "Large water drop")), y=Frequency_Hz)) + #add the axis
  geom_boxplot(aes(color=Species), outlier.shape = NA, fill = NA) +
  geom_point(aes(fill=Species), position = position_jitterdodge(.2), size = 1, color="black", pch=21) +
  labs(y="Frequency (Hz)", x=element_blank()) + #creat labels... 
  theme_classic(base_size=8) +
  theme(legend.text = element_text(size = 8, face = "italic"), legend.title = element_blank(),  legend.background = element_blank(), axis.text = element_text(color="black", size = 8), axis.ticks = element_line(color = "black"), strip.text = element_text(size = 8)) +
  theme(legend.title.align=0.5, axis.text.x = element_blank(), legend.position = c(.85, .95)) +
  scale_y_continuous(breaks= pretty_breaks(n=7)) +
  scale_color_manual(values = c("red3", "darkblue")) +
  scale_fill_manual(values = c("red3", "darkblue")) +
  facet_grid(cols = vars(Excitation_Test_Type), as.table = T, scales = "free", space = "free")
frequency_raw
```

horizontal viz means across types
```{r}
frequency_raw_perclutch_horz= ggplot(biomech_summary1.5, aes(x=Excitation_Test_Type, y=mean)) + #add the axis
  geom_boxplot(aes(color=Species), outlier.shape = NA, fill = NA) +
  geom_point(aes(fill=Species, shape=Excitation_Test_Type), position = position_jitterdodge(.2), size = 2, color="black") +
  labs(y="Frequency (Hz)", x=element_blank()) + #creat labels... 
  theme_classic(base_size=8) +
  theme(legend.text = element_text(size = 8, face = "italic"), legend.title = element_blank(),  legend.background = element_blank(), axis.text = element_text(color="black", size = 8), axis.ticks = element_line(color = "black"), axis.text.y = element_blank(), strip.text = element_text(size = 8)) +
  theme(legend.title.align=0.5, legend.position = c(.85, .95)) +
  scale_y_continuous(limits=c(0,450), breaks= pretty_breaks(n=7)) +
  scale_color_manual(values = c("red3", "darkblue")) +
  scale_fill_manual(values = c("red3", "darkblue")) +
  scale_shape_manual(values = c(24, 21)) +
  coord_flip()
frequency_raw_perclutch_horz
```

#### STATS
GLMM
```{r}
library(lme4)
library(car)
library(lmerTest)

hist(biomech$Frequency_Hz)

#frequency
modal_frequency_fullmodel = lmer(Frequency_Hz~Species*Excitation_Test +  (1|Clutch), data=biomech)
modal_frequency_nointeraction = lmer(Frequency_Hz~Species+Excitation_Test +  (1|Clutch), data=biomech)
modal_frequency_nospecies = lmer(Frequency_Hz~Excitation_Test +  (1|Clutch), data=biomech)
modal_frequency_noexcitationtest = lmer(Frequency_Hz~Species +  (1|Clutch), data=biomech)

modal_frequency_fullmodel2 = lmer(Frequency_Hz~Species*Excitation_Test*Clutch_Size +  (1|Clutch), data=biomech)
modal_frequency_fullmodel3 = lm(Frequency_Hz~Species*Excitation_Test, data=biomech)

#AIC model comparison
library(AICcmodavg)
print(aictab(list(modal_frequency_fullmodel2, modal_frequency_fullmodel,  modal_frequency_nointeraction, modal_frequency_nospecies, modal_frequency_noexcitationtest), c("modal_frequency_fullmodel2", "modal_frequency_fullmodel", "modal_frequency_nointeraction", "modal_frequency_nospecies", "modal_frequency_noexcitationtest")), LL = FALSE)

plot(fitted(modal_frequency_fullmodel), residuals(modal_frequency_fullmodel))
hist(residuals(modal_frequency_fullmodel))
qqnorm(residuals(modal_frequency_fullmodel))
qqline(residuals(modal_frequency_fullmodel))
shapiro.test(residuals(modal_frequency_fullmodel))

summary(modal_frequency_fullmodel)
Anova(modal_frequency_fullmodel)
```

Analysis of Deviance Table (Type II Wald chisquare tests)

Response: Frequency_Hz
                          Chisq Df Pr(>Chisq)    
Species                 849.529  1  < 2.2e-16 ***
Excitation_Test          60.720  3  4.125e-13 ***
Species:Excitation_Test  18.333  3  0.0003755 ***


Likelihood ratio tests
```{r}
anova(modal_frequency_fullmodel, modal_frequency_nointeraction, test = "Chisq")
anova(modal_frequency_fullmodel, modal_frequency_nospecies, test = "Chisq")
anova(modal_frequency_fullmodel, modal_frequency_noexcitationtest, test = "Chisq")

#clutch effects
anova(modal_frequency_fullmodel, modal_frequency_fullmodel3, test = "Chisq")
```

EFFECT OF INTERACTION:
* **X = 18.289; p-value = 0.0003835 **

EFFECT OF SPECIES:
* **X = 514.28; p-value < 2.2e-16 **

EFFECT OF EXCITATION TEST:
* **X = 74.58; p-value = 4.683e-14 **


##### Posthoc analyses
```{r}
library(emmeans)
# within excitation type comparisons
emms<-emmeans(modal_frequency_fullmodel, ~Species|Excitation_Test) #, adjust="Bonferoni"
pairs(emms, interaction = "pairwise") %>% rbind(adjust="fdr")

#within species comparisons
emms<-emmeans(modal_frequency_fullmodel, ~Excitation_Test|Species) #, adjust="Bonferoni"
pairs(emms, interaction = "pairwise") %>% rbind(adjust="fdr")
```

*Within Excitation Type Comparisons*
 Excitation_Test   Species_pairwise             estimate   SE  df t.ratio p.value
 Large water drop  A. callidryas - A. spurrelli    -86.6 4.79 503 -18.076  <.0001
 Pendulum impact 1 A. callidryas - A. spurrelli    -64.6 4.71 505 -13.726  <.0001
 Pendulum impact 2 A. callidryas - A. spurrelli    -66.2 4.90 510 -13.489  <.0001
 Small water drop  A. callidryas - A. spurrelli    -84.2 4.94 506 -17.042  <.0001

Degrees-of-freedom method: kenward-roger 
P value adjustment: fdr method for 4 tests 


*Within Species Comparisons*
 Species       Excitation_Test_pairwise              estimate   SE  df t.ratio p.value
 A. callidryas Large water drop - Pendulum impact 1     9.436 4.20 497   2.247  0.0502
 A. callidryas Large water drop - Pendulum impact 2     8.160 4.43 497   1.842  0.0991
 A. callidryas Large water drop - Small water drop     -1.357 4.27 495  -0.318  0.8465
 A. callidryas Pendulum impact 1 - Pendulum impact 2   -1.276 4.48 499  -0.285  0.8465
 A. callidryas Pendulum impact 1 - Small water drop   -10.794 4.31 496  -2.505  0.0301
 A. callidryas Pendulum impact 2 - Small water drop    -9.518 4.54 497  -2.096  0.0628
 
 A. spurrelli  Large water drop - Pendulum impact 1    31.413 5.02 496   6.258  <.0001
 A. spurrelli  Large water drop - Pendulum impact 2    28.567 5.05 505   5.661  <.0001
 A. spurrelli  Large water drop - Small water drop      0.985 5.14 497   0.192  0.8481
 A. spurrelli  Pendulum impact 1 - Pendulum impact 2   -2.846 4.89 504  -0.582  0.7481
 A. spurrelli  Pendulum impact 1 - Small water drop   -30.428 5.00 495  -6.086  <.0001
 A. spurrelli  Pendulum impact 2 - Small water drop   -27.582 5.04 505  -5.477  <.0001

Degrees-of-freedom method: kenward-roger 
P value adjustment: fdr method for 12 tests 



*****************************************************************************************************************
*****************************************************************************************************************
*****************************************************************************************************************




## Vibration decay over time

#### Time constant - Decay rate
```{r}
time_constant_raw= ggplot(biomech, aes(x=factor(Excitation_Test, level = c("Pendulum impact", "Large water drop", "Small water drop")), y=Time_Constant_s)) + #add the axis
  geom_point(aes(color=Species), position = position_jitterdodge(.2), size = 1) +
  geom_boxplot(aes(color=Species), outlier.shape = NA, fill = NA) +
  labs(y="Time constant (s)", x=element_blank()) + #creat labels... 
  theme_classic(base_size=8) +
  theme(legend.text = element_text(size = 8, face = "italic"), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 8), axis.ticks = element_line(color = "black")) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= c(.85,.95)) +
  scale_y_continuous(breaks= pretty_breaks(n=6)) + # allows for making "pretty" breaks on y axis... shows 5 breaks
  scale_color_manual(values = c("red3", "darkblue")) 
time_constant_raw

#Clutch Morphometrics
timeconstant_vs_clutchsize_raw= ggplot(biomech, aes(x=Clutch_Size, y=Time_Constant_s)) + #add the axis
  geom_smooth(aes(color=Species, group=interaction(Species,Excitation_Test), linetype=Species), method = "glm", se=F, size = 1 ) +
  geom_point(aes(fill=Species), position = position_jitterdodge(.2), size = 1, color="black", pch=21) +
  labs(y="Time constant (s)", x="Clutch size") + #creat labels... 
  theme_classic(base_size=8) +
  theme(legend.text = element_text(size = 8), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 8), axis.ticks = element_line(color = "black")) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= c(.025,.95)) +
  scale_y_continuous(breaks= pretty_breaks(n=6)) + # allows for making "pretty" breaks on y axis... shows 5 breaks
  scale_color_manual(values = c("red3", "darkblue")) +
  scale_fill_manual(values = c("red3", "darkblue")) 
timeconstant_vs_clutchsize_raw

timeconstant_vs_clutchsize_raw= ggplot(biomech, aes(x=Clutch_Thickness_mm, y=Time_Constant_s)) + #add the axis
  geom_smooth(aes(color=Species, group=interaction(Species,Excitation_Test), linetype=Species), method = "glm", se=F, size = 1 ) +
  geom_point(aes(fill=Species), position = position_jitterdodge(.2), size = 1, color="black", pch=21) +
  labs(y="Time constant (s)", x="Clutch Thickness") + #creat labels... 
  theme_classic(base_size=8) +
  theme(legend.text = element_text(size = 8), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 8), axis.ticks = element_line(color = "black")) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= c(.025,.95)) +
  scale_y_continuous(breaks= pretty_breaks(n=6)) + # allows for making "pretty" breaks on y axis... shows 5 breaks
  scale_color_manual(values = c("red3", "darkblue")) +
  scale_fill_manual(values = c("red3", "darkblue")) 
timeconstant_vs_clutchsize_raw

# effect of clutch size?
m2 = lmer(Time_Constant_s~Species*Clutch_Size +  (1|Clutch), data=biomech)
Anova(m2)

m2 = lmer(Time_Constant_s~Species*Clutch_Thickness_mm +  (1|Clutch), data=biomech)
Anova(m2)
```

Summarize data
```{r}
library(dplyr)
biomech_summary5 = biomech %>%
  group_by(Species, Excitation_Test, Excitation_Test_Type, Clutch) %>%
summarise(mean = mean(Time_Constant_s, na.rm=T), SE= sd(Time_Constant_s, na.rm=T)/sqrt(length(Time_Constant_s)), SD = sd(Time_Constant_s, na.rm=T), max = max(Time_Constant_s, na.rm=T), min = min(Time_Constant_s, na.rm=T), sample_size_unique=n_distinct(Time_Constant_s, na.rm=T), sample_size=n())
biomech_summary5

biomech_summary5.5 = biomech %>%
  group_by(Species, Excitation_Test) %>%
summarise(mean = mean(Time_Constant_s, na.rm=T), SE= sd(Time_Constant_s, na.rm=T)/sqrt(length(Time_Constant_s)), SD = sd(Time_Constant_s, na.rm=T), max = max(Time_Constant_s, na.rm=T), min = min(Time_Constant_s, na.rm=T), sample_size_unique=n_distinct(Time_Constant_s, na.rm=T), sample_size=n())
biomech_summary5.5

biomech_summary6 = biomech %>%
  group_by(Species) %>%
summarise(mean = mean(Time_Constant_s, na.rm=T), SE= sd(Time_Constant_s, na.rm=T)/sqrt(length(Time_Constant_s)), SD = sd(Time_Constant_s, na.rm=T), max = max(Time_Constant_s, na.rm=T), min = min(Time_Constant_s, na.rm=T), sample_size_unique=n_distinct(Time_Constant_s, na.rm=T), sample_size=n())
biomech_summary6
```

##### Visualization
```{r}
library(ggplot2)

time_constant_perclutch= ggplot(biomech_summary5, aes(x=factor(Excitation_Test, level = c("Pendulum impact 1", "Pendulum impact 2", "Small water drop", "Large water drop")), y=mean)) + #add the axis
  geom_boxplot(aes(color=Species), outlier.shape = NA, fill = NA) +
  geom_point(aes(fill=Species), position = position_jitterdodge(.2), size = 1, color="black", pch=21) +
  labs(y="Time constant (s)", x=element_blank()) + #creat labels... 
  theme_classic(base_size=8) +
  theme(legend.text = element_text(size = 8), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 8), axis.ticks = element_line(color = "black")) +
  theme(legend.title.align=0.5, legend.position = "none", strip.text = element_blank()) +
  scale_y_continuous(breaks= pretty_breaks(n=7)) + # allows for making "pretty" breaks on y axis... shows 5 breaks
  scale_color_manual(values = c("red3", "darkblue")) +
  scale_fill_manual(values = c("red3", "darkblue")) +
  facet_grid(cols = vars(Excitation_Test_Type), as.table = T, scales = "free", space = "free")
time_constant_perclutch

time_constant_perclutch= ggplot(biomech, aes(x=factor(Excitation_Test, level = c("Pendulum impact 1", "Pendulum impact 2", "Small water drop", "Large water drop")), y=Time_Constant_s)) + #add the axis
  geom_boxplot(aes(color=Species), outlier.shape = NA, fill = NA) +
  geom_point(aes(fill=Species), position = position_jitterdodge(.2), size = 1, color="black", pch=21) +
  labs(y="Time constant (s)", x=element_blank()) + #creat labels... 
  theme_classic(base_size=8) +
  theme(legend.text = element_text(size = 8), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 8), axis.ticks = element_line(color = "black")) +
  theme(legend.title.align=0.5, legend.position = "none", strip.text = element_blank()) +
  scale_y_continuous(breaks= pretty_breaks(n=7)) + # allows for making "pretty" breaks on y axis... shows 5 breaks
  scale_color_manual(values = c("red3", "darkblue")) +
  scale_fill_manual(values = c("red3", "darkblue")) +
  facet_grid(cols = vars(Excitation_Test_Type), as.table = T, scales = "free", space = "free")
time_constant_perclutch
```

##### STATS

GLMM
```{r}
library(lme4)
library(car)
library(lmerTest)

hist(biomech$Time_Constant_s)

#frequency
time_constant_fullmodel = lmer(Time_Constant_s~Species*Excitation_Test +  (1|Clutch), data=biomech)
time_constant_nointeraction = lmer(Time_Constant_s~Species+Excitation_Test +  (1|Clutch), data=biomech)
time_constant_nospecies = lmer(Time_Constant_s~Excitation_Test +  (1|Clutch), data=biomech)
time_constant_noexcitationtest = lmer(Time_Constant_s~Species +  (1|Clutch), data=biomech)

time_constant_fullmodel2 = lmer(Time_Constant_s~Species*Excitation_Test*Clutch_Size +  (1|Clutch), data=biomech)


#AIC model comparison
library(AICcmodavg)
print(aictab(list(time_constant_fullmodel2, time_constant_fullmodel,  time_constant_nointeraction, time_constant_nospecies, time_constant_noexcitationtest), c("time_constant_fullmodel2", "time_constant_fullmodel", "time_constant_nointeraction", "time_constant_nospecies", "time_constant_noexcitationtest")), LL = FALSE)

plot(fitted(time_constant_fullmodel), residuals(time_constant_fullmodel))
hist(residuals(time_constant_fullmodel))
qqnorm(residuals(time_constant_fullmodel))
qqline(residuals(time_constant_fullmodel))
shapiro.test(residuals(time_constant_fullmodel))

summary(time_constant_fullmodel)
Anova(time_constant_fullmodel)
```

Analysis of Deviance Table (Type II Wald chisquare tests)

Response: Time_Constant_s
                          Chisq Df Pr(>Chisq)    
Species                 309.911  1  < 2.2e-16 ***
Excitation_Test          10.653  3   0.013756 *  
Species:Excitation_Test  14.206  3   0.002638 ** 


Likelihood ratio tests
```{r}
anova(time_constant_fullmodel, time_constant_nointeraction, test = "Chisq")
anova(time_constant_fullmodel, time_constant_nospecies, test = "Chisq")
anova(time_constant_fullmodel, time_constant_noexcitationtest, test = "Chisq")
```

EFFECT OF INTERACTION:
* **X = 14.147; p-value = 0.002712 **

EFFECT OF SPECIES:
* **X = 255.13; p-value < 2.2e-16 **

EFFECT OF EXCITATION TEST:
* **X = 24.433; p-value = 0.0004346 **



##### Posthoc analyses
```{r}
library(emmeans)
# within excitation type comparisons
emms<-emmeans(time_constant_fullmodel, ~Species|Excitation_Test) #, adjust="Bonferoni"
pairs(emms, interaction = "pairwise") %>% rbind(adjust="fdr")

#within species comparisons
emms<-emmeans(time_constant_fullmodel, ~Excitation_Test|Species) #, adjust="Bonferoni"
pairs(emms, interaction = "pairwise") %>% rbind(adjust="fdr")
```

*Within Excitation Type Comparisons*
Excitation_Test   Species_pairwise             estimate      SE  df t.ratio p.value
 Large water drop  A. callidryas - A. spurrelli   0.0764 0.00750 502  10.192  <.0001
 Pendulum impact 1 A. callidryas - A. spurrelli   0.0484 0.00737 503   6.573  <.0001
 Pendulum impact 2 A. callidryas - A. spurrelli   0.0799 0.00769 509  10.398  <.0001
 Small water drop  A. callidryas - A. spurrelli   0.0823 0.00774 505  10.628  <.0001

Degrees-of-freedom method: kenward-roger 
P value adjustment: fdr method for 4 tests 


*Within Species Comparisons*
 Species       Excitation_Test_pairwise              estimate      SE  df t.ratio p.value
 A. callidryas Large water drop - Pendulum impact 1   0.02355 0.00657 496   3.583  0.0015
 A. callidryas Large water drop - Pendulum impact 2  -0.00597 0.00693 496  -0.860  0.7965
 A. callidryas Large water drop - Small water drop   -0.00314 0.00669 495  -0.469  0.7965
 A. callidryas Pendulum impact 1 - Pendulum impact 2 -0.02952 0.00702 499  -4.207  0.0004
 A. callidryas Pendulum impact 1 - Small water drop  -0.02669 0.00674 496  -3.958  0.0005
 A. callidryas Pendulum impact 2 - Small water drop   0.00283 0.00711 497   0.398  0.7965
 
 A. spurrelli  Large water drop - Pendulum impact 1  -0.00444 0.00786 496  -0.565  0.7965
 A. spurrelli  Large water drop - Pendulum impact 2  -0.00246 0.00791 504  -0.311  0.7965
 A. spurrelli  Large water drop - Small water drop    0.00271 0.00804 496   0.337  0.7965
 A. spurrelli  Pendulum impact 1 - Pendulum impact 2  0.00198 0.00767 503   0.258  0.7965
 A. spurrelli  Pendulum impact 1 - Small water drop   0.00715 0.00783 495   0.914  0.7965
 A. spurrelli  Pendulum impact 2 - Small water drop   0.00517 0.00789 504   0.656  0.7965

Degrees-of-freedom method: kenward-roger 
P value adjustment: fdr method for 12 tests 






*****************************************************************************************************************
*****************************************************************************************************************
*****************************************************************************************************************


# Figure 3
```{r}
library(gridExtra)
library(cowplot)

#  @ portrait at 7.08661 x 8.26772 max for JEB
# 4x8.5 portrait looks good

Figure3 = plot_grid(frequency_raw_perclutch, time_constant_perclutch, labels = "AUTO", label_x = -.01, label_y = 1, scale =1, ncol=1, nrow = 2, label_size = 12, align = "v")
Figure3

#@ 3.46x6
```


*****************************************************************************************************************
*****************************************************************************************************************
*****************************************************************************************************************



## Peak Amplitude
```{r}
peakamplitude_summary = entire_disturbance_biomechmeasurements %>% 
  group_by(Species, Excitation_Test, Excitation_Test_Type, Clutch) %>%
summarise(mean = mean(Peak_Amplitude_G, na.rm=T), SE= sd(Peak_Amplitude_G, na.rm=T)/sqrt(length(Peak_Amplitude_G)), SD = sd(Peak_Amplitude_G, na.rm=T), max = max(Peak_Amplitude_G, na.rm=T), min = min(Peak_Amplitude_G, na.rm=T), sample_size=n(), sample_size_unique=n_distinct(Peak_Amplitude_G, na.rm=T))
peakamplitude_summary
```


##### Visualization
```{r}
peak_amplitude_raw = ggplot(entire_disturbance_biomechmeasurements, aes(x=factor(Excitation_Test, level =c("Pendulum impact 1", "Pendulum impact 2", "Small water drop", "Large water drop")), y=Peak_Amplitude_G)) + #add the axis
  geom_boxplot(aes(color=Species), fill = NA, outlier.shape = NA) + 
  geom_point(aes(fill=Species), position = position_jitterdodge(.2), size = 1, color="black", pch=21) +
  labs(y="Peak amplitude (m/s^2)", x= element_blank()) +
  theme_classic(base_size=9) +
  theme(legend.text = element_text(size = 9), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 9), axis.ticks = element_line(color = "black"), axis.text.x = element_blank(), strip.text = element_text(size = 8)) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= c(.01,.95), strip.background = element_rect(fill = NA)) +
  scale_y_continuous(breaks= pretty_breaks(n=6)) + # allows for making "pretty" breaks on y axis... shows 5 breaks
  scale_color_manual(values = c("red3", "darkblue")) +
  scale_fill_manual(values = c("red3", "darkblue")) +
  facet_grid(cols = vars(Excitation_Test_Type), as.table = T, scales = "free", space = "free")
peak_amplitude_raw
```




Clutch Morphometrics
```{r}
peak_amplitude_by_clutchsize_raw = ggplot(entire_disturbance_biomechmeasurements, aes(x=Clutch_Size, y=Peak_Amplitude_G)) + #add the axis
  geom_smooth(aes(color=Species, group=interaction(Species,Excitation_Test), linetype=Species), method = "glm", se=F, size = 1 ) +
  geom_point(aes(fill=Species), position = position_jitterdodge(.2), size = 1, color="black", pch=21) +
  labs(y="Peak amplitude (G)", x= element_blank()) +
  theme_classic(base_size=9) +
  theme(legend.text = element_text(size = 9), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 9), axis.ticks = element_line(color = "black")) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= c(.01,.95)) +
  scale_y_continuous(breaks= pretty_breaks(n=6)) + # allows for making "pretty" breaks on y axis... shows 5 breaks
  scale_color_manual(values = c("red3", "darkblue")) +
  scale_fill_manual(values = c("red3", "darkblue"))
peak_amplitude_by_clutchsize_raw

peak_amplitude_by_clutchsize_raw = ggplot(entire_disturbance_biomechmeasurements, aes(x=Clutch_Thickness_mm, y=Peak_Amplitude_G)) + #add the axis
  geom_smooth(aes(color=Species, group=interaction(Species,Excitation_Test), linetype=Species), method = "glm", se=F, size = 1 ) +
  geom_point(aes(fill=Species), position = position_jitterdodge(.2), size = 1, color="black", pch=21) +
  labs(y="Peak amplitude (G)", x= element_blank()) +
  theme_classic(base_size=9) +
  theme(legend.text = element_text(size = 9), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 9), axis.ticks = element_line(color = "black")) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= c(.01,.95)) +
  scale_y_continuous(breaks= pretty_breaks(n=6)) + # allows for making "pretty" breaks on y axis... shows 5 breaks
  scale_color_manual(values = c("red3", "darkblue")) +
  scale_fill_manual(values = c("red3", "darkblue"))
peak_amplitude_by_clutchsize_raw

# effect of clutch size?
m2 = lmer(Peak_Amplitude_G~Species*Clutch_Size*Excitation_Test +  (1|Clutch), data=entire_disturbance_biomechmeasurements)
Anova(m2)

m2 = lmer(Peak_Amplitude_G~Species*Clutch_Size +  (1|Clutch), data=entire_disturbance_biomechmeasurements)
Anova(m2)
```



##### STATS

LMM
```{r}
library(lme4)
library(car)
library(lmerTest)

hist(biomechmeasurements$Peak_Amplitude_G)

#amkplitude
peak_amplitude_fullmodel = lmer(Peak_Amplitude_G~Species*Excitation_Test +  (1|Clutch), data=entire_disturbance_biomechmeasurements)
peak_amplitude_nointeraction = lmer(Peak_Amplitude_G~Species+Excitation_Test +  (1|Clutch), data=entire_disturbance_biomechmeasurements)
peak_amplitude_nospecies = lmer(Peak_Amplitude_G~Excitation_Test +  (1|Clutch), data=entire_disturbance_biomechmeasurements)
peak_amplitude_noexcitationtest = lmer(Peak_Amplitude_G~Species +  (1|Clutch), data=entire_disturbance_biomechmeasurements)

peak_amplitude_fullmodel2 = lmer(Peak_Amplitude_G~Species*Excitation_Test*Clutch_Size +  (1|Clutch), data=entire_disturbance_biomechmeasurements)


#AIC model comparison
library(AICcmodavg)
print(aictab(list(peak_amplitude_fullmodel2, peak_amplitude_fullmodel,  peak_amplitude_nointeraction, peak_amplitude_nospecies, peak_amplitude_noexcitationtest), c("peak_amplitude_fullmodel2", "peak_amplitude_fullmodel", "peak_amplitude_nointeraction", "peak_amplitude_nospecies", "peak_amplitude_noexcitationtest")), LL = FALSE)

plot(fitted(peak_amplitude_fullmodel), residuals(peak_amplitude_fullmodel))
hist(residuals(peak_amplitude_fullmodel))
qqnorm(residuals(peak_amplitude_fullmodel))
qqline(residuals(peak_amplitude_fullmodel))
shapiro.test(residuals(peak_amplitude_fullmodel))

summary(peak_amplitude_fullmodel)
Anova(peak_amplitude_fullmodel)
```

Analysis of Deviance Table (Type II Wald chisquare tests)

Response: Peak_Amplitude_G
                         Chisq Df Pr(>Chisq)    
Species                 33.296  1  7.913e-09 ***
Excitation_Test         50.657  3  5.787e-11 ***
Species:Excitation_Test 39.045  3  1.698e-08 ***


Likelihood ratio tests
```{r}
anova(peak_amplitude_fullmodel, peak_amplitude_nointeraction, test = "Chisq")
anova(peak_amplitude_fullmodel, peak_amplitude_nospecies, test = "Chisq")
anova(peak_amplitude_fullmodel, peak_amplitude_noexcitationtest, test = "Chisq")
```

EFFECT OF INTERACTION:
* **X = 35.349; p-value = 1.028e-07 **

EFFECT OF SPECIES:
* **X = 59.464; p-value = 3.76e-12 **

EFFECT OF EXCITATION TEST:
* **X = 69.055; p-value = 6.387e-13 **



##### Posthoc analyses
```{r}
library(emmeans)
# within excitation type comparisons
emms<-emmeans(peak_amplitude_fullmodel, ~Species|Excitation_Test) #, adjust="Bonferoni"
pairs(emms, interaction = "pairwise") %>% rbind(adjust="fdr")

#within species comparisons
emms<-emmeans(peak_amplitude_fullmodel, ~Excitation_Test|Species) #, adjust="Bonferoni"
pairs(emms, interaction = "pairwise") %>% rbind(adjust="fdr")
```

*Within Excitation Type Comparisons*
 Excitation_Test   Species_pairwise             estimate    SE  df t.ratio p.value
 Large water drop  A. callidryas - A. spurrelli  0.04942 0.168 107   0.294  0.9618
 Pendulum impact 1 A. callidryas - A. spurrelli -1.05821 0.166 108  -6.388  <.0001
 Pendulum impact 2 A. callidryas - A. spurrelli -0.95910 0.168 109  -5.693  <.0001
 Small water drop  A. callidryas - A. spurrelli  0.00821 0.171 106   0.048  0.9618

Degrees-of-freedom method: kenward-roger 
P value adjustment: fdr method for 4 tests 


*Within Species Comparisons*
 Species       Excitation_Test_pairwise              estimate    SE  df t.ratio p.value
 A. callidryas Large water drop - Pendulum impact 1    0.0664 0.161 102   0.412  0.6811
 A. callidryas Large water drop - Pendulum impact 2    0.4911 0.164 102   3.000  0.0068
 A. callidryas Large water drop - Small water drop     0.4015 0.161 102   2.494  0.0213
 A. callidryas Pendulum impact 1 - Pendulum impact 2   0.4247 0.164 102   2.594  0.0186
 A. callidryas Pendulum impact 1 - Small water drop    0.3352 0.161 102   2.082  0.0501
 A. callidryas Pendulum impact 2 - Small water drop   -0.0895 0.164 102  -0.547  0.6389
 
 A. spurrelli  Large water drop - Pendulum impact 1   -1.0413 0.169 103  -6.160  <.0001
 A. spurrelli  Large water drop - Pendulum impact 2   -0.5175 0.169 103  -3.061  0.0067
 A. spurrelli  Large water drop - Small water drop     0.3603 0.175 103   2.061  0.0501
 A. spurrelli  Pendulum impact 1 - Pendulum impact 2   0.5238 0.166 102   3.157  0.0063
 A. spurrelli  Pendulum impact 1 - Small water drop    1.4016 0.172 104   8.125  <.0001
 A. spurrelli  Pendulum impact 2 - Small water drop    0.8778 0.172 104   5.089  <.0001

Degrees-of-freedom method: kenward-roger 
P value adjustment: fdr method for 12 tests 


*****************************************************************************************************************
*****************************************************************************************************************
*****************************************************************************************************************





## Vibration Duration

```{r}
#Clutch Morphometrics
timeconstant_vs_clutchsize_raw= ggplot(biomechmeasurements, aes(x=Clutch_Size, y=Mean_Duration_s)) + #add the axis
  geom_smooth(aes(color=Species, group=interaction(Species,Excitation_Test), linetype=Species), method = "glm", se=F, size = 1 ) +
  geom_point(aes(fill=Species), position = position_jitterdodge(.2), size = 1, color="black", pch=21) +
  labs(y="Vibration duratino (s)", x="Clutch size") + #creat labels... 
  theme_classic(base_size=8) +
  theme(legend.text = element_text(size = 8), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 8), axis.ticks = element_line(color = "black")) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= c(.025,.95)) +
  scale_y_continuous(breaks= pretty_breaks(n=6)) + # allows for making "pretty" breaks on y axis... shows 5 breaks
  scale_color_manual(values = c("red3", "darkblue")) +
  scale_fill_manual(values = c("red3", "darkblue")) 
timeconstant_vs_clutchsize_raw

timeconstant_vs_clutchsize_raw= ggplot(biomechmeasurements, aes(x=Clutch_Thickness_mm, y=Mean_Duration_s)) + #add the axis
  geom_smooth(aes(color=Species, group=interaction(Species,Excitation_Test), linetype=Species), method = "glm", se=F, size = 1 ) +
  geom_point(aes(fill=Species), position = position_jitterdodge(.2), size = 1, color="black", pch=21) +
  labs(y="Vibratin duration (s)", x="Clutch Thickness") + #creat labels... 
  theme_classic(base_size=8) +
  theme(legend.text = element_text(size = 8), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 8), axis.ticks = element_line(color = "black")) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= c(.025,.95)) +
  scale_y_continuous(breaks= pretty_breaks(n=6)) + # allows for making "pretty" breaks on y axis... shows 5 breaks
  scale_color_manual(values = c("red3", "darkblue")) +
  scale_fill_manual(values = c("red3", "darkblue")) 
timeconstant_vs_clutchsize_raw

# effect of clutch size?
m2 = lmer(Mean_Duration_s~Species*Clutch_Size*Excitation_Test +  (1|Clutch), data=biomechmeasurements)
Anova(m2)

m2 = lmer(Mean_Duration_s~Species*Clutch_Size +  (1|Clutch), data=biomechmeasurements)
Anova(m2)
```


```{r}
vibrationduration_summary = entire_disturbance_biomechmeasurements %>% 
  group_by(Species, Excitation_Test, Excitation_Test_Type, Clutch) %>%
summarise(mean = mean(Mean_Duration_s, na.rm=T), SE= sd(Mean_Duration_s, na.rm=T)/sqrt(length(Mean_Duration_s)), SD = sd(Mean_Duration_s, na.rm=T), max = max(Mean_Duration_s, na.rm=T), min = min(Mean_Duration_s, na.rm=T), sample_size=n(), sample_size_unique=n_distinct(Mean_Duration_s, na.rm=T))
vibrationduration_summary

vibrationduration_summary2 = entire_disturbance_biomechmeasurements %>% 
  group_by(Species) %>%
summarise(mean = mean(Mean_Duration_s, na.rm=T), SE= sd(Mean_Duration_s, na.rm=T)/sqrt(length(Mean_Duration_s)), SD = sd(Mean_Duration_s, na.rm=T), max = max(Mean_Duration_s, na.rm=T), min = min(Mean_Duration_s, na.rm=T), sample_size=n(), sample_size_unique=n_distinct(Mean_Duration_s, na.rm=T))
vibrationduration_summary2

vibrationduration_summary3 = entire_disturbance_biomechmeasurements %>% 
  group_by(Excitation_Test_Type) %>%
summarise(mean = mean(Mean_Duration_s, na.rm=T), SE= sd(Mean_Duration_s, na.rm=T)/sqrt(length(Mean_Duration_s)), SD = sd(Mean_Duration_s, na.rm=T), max = max(Mean_Duration_s, na.rm=T), min = min(Mean_Duration_s, na.rm=T), sample_size=n(), sample_size_unique=n_distinct(Mean_Duration_s, na.rm=T))
vibrationduration_summary3
```


##### Visualization
```{r}
vibration_duration_plot2 = ggplot(entire_disturbance_biomechmeasurements, aes(x=factor(Excitation_Test, level =c("Pendulum impact 1", "Pendulum impact 2", "Small water drop", "Large water drop")), y=Mean_Duration_s)) + #add the axis
  geom_boxplot(aes(color=Species), fill = NA, outlier.shape = NA) + 
  geom_point(aes(fill=Species), position = position_jitterdodge(),  size = 1, color = "black", pch=21) +
  labs(y="Vibration duration (s)", x= element_blank()) +
  theme_classic(base_size=9) +
  theme(legend.text = element_text(size = 9), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 9), axis.ticks = element_line(color = "black"), axis.text.x = element_blank()) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= "none", strip.text = element_text(size = 8), strip.background = element_rect(fill = NA)) +
  scale_y_continuous(breaks= pretty_breaks(n=7)) + # allows for making "pretty" breaks on y axis... shows 5 breaks
  scale_color_manual(values = c("red3", "darkblue")) +
  scale_fill_manual(values = c("red3", "darkblue")) +
  facet_grid(cols = vars(Excitation_Test_Type), as.table = T, scales = "free", space = "free")
vibration_duration_plot2

vibration_duration_plot2 = ggplot(entire_disturbance_biomechmeasurements, aes(x=factor(Excitation_Test, level =c("Pendulum impact 1", "Pendulum impact 2", "Small water drop", "Large water drop")), y=Mean_Duration_s)) + #add the axis
  geom_boxplot(aes(color=Species), fill = NA, outlier.shape = NA) + 
  geom_point(aes(fill=Species), position = position_jitterdodge(),  size = 1, color = "black", pch=21) +
  labs(y="Vibration duration (s)", x= element_blank()) +
  theme_classic(base_size=9) +
  theme(legend.text = element_text(size = 9), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 9), axis.ticks = element_line(color = "black")) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= "none", strip.text = element_blank())+
  scale_y_continuous(breaks= pretty_breaks(n=7)) + # allows for making "pretty" breaks on y axis... shows 5 breaks
  scale_color_manual(values = c("red3", "darkblue")) +
  scale_fill_manual(values = c("red3", "darkblue")) +
  facet_grid(cols = vars(Excitation_Test_Type), as.table = T, scales = "free", space = "free")
vibration_duration_plot2
```

##### STATS

LMM
```{r}
library(lme4)
library(car)
library(lmerTest)

hist(entire_disturbance_biomechmeasurements$Mean_Duration_s)

#frequency
vibration_duration_fullmodel = lmer(Mean_Duration_s~Species*Excitation_Test +  (1|Clutch), data=entire_disturbance_biomechmeasurements)
vibration_duration_nointeraction = lmer(Mean_Duration_s~Species+Excitation_Test +  (1|Clutch), data=entire_disturbance_biomechmeasurements)
vibration_duration_nospecies = lmer(Mean_Duration_s~Excitation_Test +  (1|Clutch), data=entire_disturbance_biomechmeasurements)
vibration_duration_noexcitationtest = lmer(Mean_Duration_s~Species +  (1|Clutch), data=entire_disturbance_biomechmeasurements)

vibration_duration_fullmodel2 = lmer(Mean_Duration_s~Species*Excitation_Test*Clutch_Size +  (1|Clutch), data=entire_disturbance_biomechmeasurements)

# model with clutch size and with impact distance is worse

#AIC model comparison
library(AICcmodavg)
print(aictab(list(vibration_duration_fullmodel2, vibration_duration_fullmodel,  vibration_duration_nointeraction, vibration_duration_nospecies, vibration_duration_noexcitationtest), c( "vibration_duration_fullmodel2", "vibration_duration_fullmodel", "vibration_duration_nointeraction", "vibration_duration_nospecies", "vibration_duration_noexcitationtest")), LL = FALSE)

plot(fitted(vibration_duration_fullmodel), residuals(vibration_duration_fullmodel))
hist(residuals(vibration_duration_fullmodel))
qqnorm(residuals(vibration_duration_fullmodel))
qqline(residuals(vibration_duration_fullmodel))
shapiro.test(residuals(vibration_duration_fullmodel))

summary(vibration_duration_fullmodel)
Anova(vibration_duration_fullmodel)
```

Analysis of Deviance Table (Type II Wald chisquare tests)

Response: Mean_Duration_s
                           Chisq Df Pr(>Chisq)    
Species                 233.6487  1  < 2.2e-16 ***
Excitation_Test          44.9238  3  9.604e-10 ***
Species:Excitation_Test   4.4417  3     0.2175 


Likelihood ratio tests
```{r}
anova(vibration_duration_fullmodel, vibration_duration_nointeraction, test = "Chisq")
anova(vibration_duration_fullmodel, vibration_duration_nospecies, test = "Chisq")
anova(vibration_duration_fullmodel, vibration_duration_noexcitationtest, test = "Chisq")
```

EFFECT OF INTERACTION:
* **X = 4.6407; p-value = 0.2001 **

EFFECT OF SPECIES:
* **X = 136.96; p-value < 2.2e-16 **

EFFECT OF EXCITATION TEST:
* **X = 42.988; p-value = 1.173e-07 **



##### Posthoc analyses
```{r}
library(emmeans)
# within excitation type comparisons
emms<-emmeans(vibration_duration_fullmodel, ~Species|Excitation_Test) #, adjust="Bonferoni"
pairs(emms, interaction = "pairwise") %>% rbind(adjust="fdr")

#within species comparisons
emms<-emmeans(vibration_duration_fullmodel, ~Excitation_Test|Species) #, adjust="Bonferoni"
pairs(emms, interaction = "pairwise") %>% rbind(adjust="fdr")
```

*Within Excitation Type Comparisons*
 Excitation_Test   Species_pairwise             estimate     SE  df t.ratio p.value
 Large water drop  A. callidryas - A. spurrelli    0.270 0.0365 105   7.386  <.0001
 Pendulum impact 1 A. callidryas - A. spurrelli    0.339 0.0361 106   9.399  <.0001
 Pendulum impact 2 A. callidryas - A. spurrelli    0.316 0.0367 107   8.619  <.0001
 Small water drop  A. callidryas - A. spurrelli    0.242 0.0371 104   6.534  <.0001

Degrees-of-freedom method: kenward-roger 
P value adjustment: fdr method for 4 tests 


*Within Species Comparisons*
 Species       Excitation_Test_pairwise              estimate     SE  df t.ratio p.value
 A. callidryas Large water drop - Pendulum impact 1  -0.15606 0.0347 101  -4.500  0.0001
 A. callidryas Large water drop - Pendulum impact 2  -0.10505 0.0353 102  -2.977  0.0109
 A. callidryas Large water drop - Small water drop    0.03696 0.0347 101   1.066  0.3468
 A. callidryas Pendulum impact 1 - Pendulum impact 2  0.05101 0.0353 102   1.446  0.2018
 A. callidryas Pendulum impact 1 - Small water drop   0.19302 0.0347 101   5.566  <.0001
 A. callidryas Pendulum impact 2 - Small water drop   0.14202 0.0353 102   4.025  0.0004
 
 A. spurrelli  Large water drop - Pendulum impact 1  -0.08686 0.0365 102  -2.380  0.0383
 A. spurrelli  Large water drop - Pendulum impact 2  -0.05848 0.0365 102  -1.602  0.1682
 A. spurrelli  Large water drop - Small water drop    0.00941 0.0377 102   0.249  0.8035
 A. spurrelli  Pendulum impact 1 - Pendulum impact 2  0.02838 0.0357 101   0.794  0.4682
 A. spurrelli  Pendulum impact 1 - Small water drop   0.09626 0.0373 103   2.581  0.0270
 A. spurrelli  Pendulum impact 2 - Small water drop   0.06789 0.0373 103   1.820  0.1228

Degrees-of-freedom method: kenward-roger 
P value adjustment: fdr method for 12 tests 



*****************************************************************************************************************
*****************************************************************************************************************
*****************************************************************************************************************





# Figure 4
```{r}
library(gridExtra)
library(cowplot)

#  @ portrait at 7.08661 x 8.26772 max for JEB
# 4x8.5 portrait looks good

Figure4 = plot_grid(peak_amplitude_raw, vibration_duration_plot2, labels = "AUTO", label_x = -.01, label_y = 1, scale =1, ncol=1, nrow = 2, label_size = 12, align = "v")
Figure4

#@ 3.46x6
```



*****************************************************************************************************************
*****************************************************************************************************************
*****************************************************************************************************************


*****************************************************************************************************************
*****************************************************************************************************************
*****************************************************************************************************************




### Amplitude Loss

```{r}
amploss = entire_disturbance_biomechmeasurements %>%
  filter(Excitation_Test_Type == "Pendulum impact")
amploss

amploss_summary = amploss %>% 
  group_by(Species, Excitation_Test_Subtype, Vibration_Signal_Portion) %>%
summarise(mean = mean(Peak_Amplitude_G, na.rm=T), SE= sd(Peak_Amplitude_G, na.rm=T)/sqrt(length(Peak_Amplitude_G)), SD = sd(Peak_Amplitude_G, na.rm=T), max = max(Peak_Amplitude_G, na.rm=T), min = min(Peak_Amplitude_G, na.rm=T), sample_size=n(), sample_size_unique=n_distinct(Peak_Amplitude_G, na.rm=T))
amploss_summary
```


##### Visualization
```{r}
library(lemon)
amplitude_loss_plot1 = ggplot(amploss, aes(x=as.character(Excitation_Test_Subtype), y=Peak_Amplitude_G)) + #add the axis
  lemon::geom_pointpath(aes(fill=Species, group=Clutch), size=1,  shorten = F, distance=0, color="black", pch=21) +
  labs(y="Peak amplitude (m/s^2)", x= "Distance to accelerometer (cm)") +
  theme_classic(base_size=9) +
  theme(legend.text = element_text(size = 9), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 9), axis.ticks = element_line(color = "black"), strip.text = element_blank(), strip.placement = "none") +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= "none") +
  scale_y_continuous(breaks= pretty_breaks(n=6)) + # allows for making "pretty" breaks on y axis... shows 5 breaks
  scale_color_manual(values = c("red3", "darkblue")) +
  scale_fill_manual(values = c("red3", "darkblue")) +
  facet_rep_grid(col=vars(Species))
amplitude_loss_plot1
```


### Amplitude difference - proportion

Import data from entire disturbance
```{r}
library(readxl)
amploss_props <- read_excel("~/Desktop/Publications/Manuscripts/In Prep/Comparative Egg Clutch Biomechanics/Comparative Egg Clutch Biomechanics.xlsx", 
    sheet = "Amplitude Comparisons", na = "NA")
amploss_props
```

##### Visualization
```{r}
library(lemon)
library(ggplot2)
amplitude_loss_prop_plot1 = ggplot(amploss_props, aes(x=Species, y=Proportion_Amplitude)) + #add the axis
  geom_hline(yintercept=1, linetype="dashed", color = "black", size=.5)+
  geom_boxplot(aes(color=Species, group=interaction(Species,Comparison)), fill = NA, outlier.shape = NA) + 
  geom_point(aes(fill=Species), position = position_jitterdodge(.3), size = 1, color = "black", pch=21) +
  labs(y="Proportion of peak amplitude", x= element_blank()) +
  theme_classic(base_size=9) +
  theme(legend.text = element_text(size = 9), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 9), axis.ticks = element_line(color = "black"), axis.text.x = element_text(face = "italic")) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= "none") +
  scale_y_continuous(breaks= pretty_breaks(n=6)) +
  scale_fill_manual(values = c("red3", "darkblue")) +
  scale_color_manual(values = c("red3", "darkblue")) 
amplitude_loss_prop_plot1
```


#### STATS
stats on raw difference
```{r}
hist(amploss_props$Raw_Difference)

m=lm(Raw_Difference~Species, data = amploss_props)

plot(fitted(m), residuals(m))
hist(residuals(m))
qqnorm(residuals(m))
qqline(residuals(m))
shapiro.test(residuals(m))

t.test(Raw_Difference~Species, data = amploss_props)
```

Welch Two Sample t-test

data:  Raw_Difference by Species
t = -0.94771, df = 18.79, p-value = 0.3553
alternative hypothesis: true difference in means between group A. callidryas and group A. spurrelli is not equal to 0
95 percent confidence interval:
 -0.6275243  0.2365660
sample estimates:
mean in group A. callidryas  mean in group A. spurrelli 
                  0.3283333                   0.5238125 
                  
stats on proportion data
```{r}
hist(amploss_props$Proportion_Amplitude)

m=lm(Proportion_Amplitude~Species, data = amploss_props)

plot(fitted(m), residuals(m))
hist(residuals(m))
qqnorm(residuals(m))
qqline(residuals(m))
shapiro.test(residuals(m))

t.test(Proportion_Amplitude~Species, data = amploss_props)

# exclude values of proportion peak amplitude over 1
amploss_props_exlcude = amploss_props %>%
  filter(Proportion_Amplitude<1)
t.test(Proportion_Amplitude~Species, data = amploss_props_exlcude)
```

Shapiro-Wilk normality test

data:  residuals(m)
W = 0.97169, p-value = 0.5668

Welch Two Sample t-test

data:  Proportion_Amplitude by Species
*t = -1.6567, df = 26.578, p-value = 0.1093*
alternative hypothesis: true difference in means between group A. callidryas and group A. spurrelli is not equal to 0
95 percent confidence interval:
 -0.40017789  0.04278954
sample estimates:
mean in group A. callidryas  mean in group A. spurrelli 
                  0.5833806                   0.7620748 


# Figure 5
```{r}
library(gridExtra)
library(cowplot)

#  @ portrait at 7.08661 x 8.26772 max for JEB
# 4x8.5 portrait looks good

Figure5 = plot_grid(amplitude_loss_plot1, amplitude_loss_prop_plot1, labels = "AUTO", label_x = -.01, label_y = 1, scale =1, ncol=1, nrow = 2, label_size = 12, align = "v")
Figure5
# @3.46 x 6
```



*****************************************************************************************************************




# Egg-Transplant Egg-Predation Trials
```{r}
library(readxl)
transplant <- read_excel("~/Desktop/Publications/Manuscripts/In Prep/Comparative Egg Clutch Biomechanics/Comparative Egg Clutch Biomechanics.xlsx", 
    sheet = "Egg-Transplant Predation Trials", na = "NA")
transplant
```

Summarize only by treatment
```{r}
library(dplyr)
transplant_summary = transplant %>% 
  group_by(Species,Treatment) %>%
summarise(mean_escape_success = mean(Escape_Success, na.rm=T), SE= sd(Escape_Success, na.rm=T)/sqrt(length(Escape_Success)), SD = sd(Escape_Success, na.rm=T), min= min(Escape_Success, na.rm=T), max= max(Escape_Success, na.rm=T), sample_size=n())
transplant_summary

#Number available to hatch per species
transplant_summary2 = transplant %>% 
  group_by(Species,Treatment) %>%
summarise(mean_avail_to_hatch = mean(Number_Available_to_Hatch, na.rm=T), SE= sd(Number_Available_to_Hatch, na.rm=T)/sqrt(length(Number_Available_to_Hatch)), SD = sd(Number_Available_to_Hatch, na.rm=T), min= min(Number_Available_to_Hatch, na.rm=T), max= max(Number_Available_to_Hatch, na.rm=T), sample_size=n())
transplant_summary2
```




## Figure 6

Escape Success Summarized Reaction Norm
```{r}
library(ggplot2)
library(scales)

#This shows reaction norms across clutch effects —i.e. sibship effect

# Figure S2

mean_escape_success = ggplot(transplant, aes(x=Treatment, y=Escape_Success)) + #add the axis
  geom_boxplot(aes(color=Treatment), outlier.shape = NA, fill = NA) + 
  geom_line(aes(group=Clutch), size = 1, position = position_jitter(width = 0.25, seed = 123), alpha=0.4) +
  geom_point(aes(color=Treatment, fill= Treatment), position = position_jitter(width = 0.25, seed = 123), color = "black", pch=21) +
  labs(y="Escape success (Proportion hatched)", x=element_blank()) + #create labels... 
  theme_classic(base_size=9) +
  theme(legend.text = element_text(size = 9), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 9), axis.ticks = element_line(color = "black"), strip.text = element_text(face = "italic")) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= "none") +
  scale_y_continuous(breaks= pretty_breaks(n=6)) + 
  scale_color_manual(values = c("darkgreen", "darkgreen","darkorange3")) +
  scale_fill_manual(values = c("darkgreen", "darkgreen","darkorange3")) +
  geom_text(aes(y= .9, x = 1.5, label= "***"), color= "gray", vjust = 0, size = 4) + #add significance star/astrix 
  geom_segment(aes(x = 1, y = .9, xend = 2, yend = .9), color="gray", size = .25) +# add line segment 
  facet_grid(~factor(Species, levels=c('A. spurrelli','A. callidryas')), scales = "free")
mean_escape_success
```

Figure 5
```{r}
#This shows reaction norms across egg mass effects —i.e. physical structure effect
transplant_sorted <-
  transplant %>%
  arrange(Egg_Mass, Treatment)

mean_escape_successV2 = ggplot(transplant, aes(x=Treatment, y=Escape_Success)) + #add the axis
  geom_boxplot(aes(color=Treatment), outlier.shape = NA, fill = NA) + 
  geom_line(data=transplant_sorted, aes(group=Egg_Mass), size = 1, position = position_jitter(width = 0.25, seed = 123), alpha=0.4) +
  geom_point(aes(color=Treatment, fill= Treatment), position = position_jitter(width = 0.25, seed = 123), color = "black", pch=21) +
  labs(y="Escape success (Proportion hatched)", x=element_blank()) + #create labels... 
  theme_classic(base_size=9) +
  theme(legend.text = element_text(size = 9), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 9), axis.ticks = element_line(color = "black"), strip.text = element_text(face = "italic")) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= "none") +
  scale_y_continuous(breaks= pretty_breaks(n=6)) + 
  scale_color_manual(values = c("darkgreen", "darkgreen","darkorange3")) +
  scale_fill_manual(values = c("darkgreen", "darkgreen","darkorange3")) +
  geom_text(aes(y= .9, x = 1.5, label= "***"), color= "gray", vjust = 0, size = 4) + #add significance star/astrix 
  geom_segment(aes(x = 1, y = .9, xend = 2, yend = .9), color="gray", size = .25) +# add line segment 
  facet_grid(~factor(Species, levels=c('A. spurrelli','A. callidryas')), scales = "free")
mean_escape_successV2

#3x3 looks good in PDF
```

## STATS

prep dataset
```{r}
transplant2 = transplant %>%
  mutate(Treatmentfortest = case_when(Treatment == "Transplanted" ~ "Transplanted", Treatment == "Non-transplanted control" ~ "Control", Treatment == "aTransplant host control" ~ "Control"))
transplant2
```



All together
```{r}
library(car)
#all models
eggpredation_full = glmer(Escape_Success~Treatmentfortest*Species +  (1|Clutch) , family="binomial", weights = Number_Available_to_Hatch, data = transplant2)
eggpredation_nointeraction = glmer(Escape_Success~Species+Treatmentfortest +  (1|Clutch), family="binomial", weights = Number_Available_to_Hatch, data = transplant2)
eggpredation_nospecies = glmer(Escape_Success~Treatmentfortest +  (1|Clutch), family="binomial", weights = Number_Available_to_Hatch, data = transplant2)
eggpredation_notreatment = glmer(Escape_Success~Species +  (1|Clutch), family="binomial", weights = Number_Available_to_Hatch, data = transplant2)


eggpredation_full2 = glmer(Escape_Success~Species*Treatmentfortest +  (1|Clutch) +(1|Egg_Mass) , family="binomial", weights = Number_Available_to_Hatch, data = transplant2)

eggpredation_full3 = glmer(Escape_Success~Species*Treatmentfortest +  (1|Egg_Mass) , family="binomial", weights = Number_Available_to_Hatch, data = transplant2)


#AIC model comparison
library(AICcmodavg)
print(aictab(list(eggpredation_full3, eggpredation_full2, eggpredation_full,  eggpredation_nointeraction, eggpredation_nospecies, eggpredation_notreatment), c("eggpredation_full3", "eggpredation_full2", "eggpredation_full", "eggpredation_nointeraction", "eggpredation_nospecies", "eggpredation_notreatment")), LL = FALSE)

plot(fitted(eggpredation_full), residuals(eggpredation_full))
hist(residuals(eggpredation_full))
qqnorm(residuals(eggpredation_full))
qqline(residuals(eggpredation_full))
shapiro.test(residuals(eggpredation_full))

summary(eggpredation_full)
Anova(eggpredation_full)
```

Analysis of Deviance Table (Type II Wald chisquare tests)

Response: Escape_Success
                           Chisq Df Pr(>Chisq)    
Species                  73.4010  1  < 2.2e-16 ***
Treatmentfortest         31.6772  1  1.821e-08 ***
Species:Treatmentfortest  5.5458  1    0.01852 *  


Likelihood ratio tests
```{r}
anova(eggpredation_full, eggpredation_nointeraction, test = "Chisq")
anova(eggpredation_full, eggpredation_nospecies, test = "Chisq")
anova(eggpredation_full, eggpredation_notreatment, test = "Chisq")
```

EFFECT OF INTERACTION:
* **X = 5.3423; p-value = 0.02081 **

EFFECT OF SPECIES:
* **X = 52.917; p-value = 3.23e-12 **

EFFECT OF TREATMENT:
* **X = 33.124; p-value = 6.416e-08 **



##### Posthoc analyses
```{r}

# within excitation type comparisons
library(emmeans)
emms=emmeans(eggpredation_full, ~Treatmentfortest|Species) #, adjust="Bonferoni"
pairs(emms, interaction = "pairwise") %>% rbind(adjust="fdr")

#within species comparisons
emms=emmeans(eggpredation_full, ~Species|Treatmentfortest) #, adjust="Bonferoni"
pairs(emms, interaction = "pairwise") %>% rbind(adjust="fdr")
```

*Within Excitation Type Comparisons*
Treatmentfortest Species_pairwise             estimate    SE  df z.ratio p.value
 Control          A. callidryas - A. spurrelli     2.93 0.338 Inf   8.672  <.0001
 Transplanted     A. callidryas - A. spurrelli     1.97 0.389 Inf   5.069  <.0001

Results are given on the log odds ratio (not the response) scale. 
P value adjustment: fdr method for 2 tests 


*Within Species Comparisons*
  Species       Treatmentfortest_pairwise estimate    SE  df z.ratio p.value
 A. callidryas Control - Transplanted      -0.656 0.294 Inf  -2.232  0.0256
 A. spurrelli  Control - Transplanted      -1.615 0.284 Inf  -5.693  <.0001

Results are given on the log odds ratio (not the response) scale. 
P value adjustment: fdr method for 2 tests 




*****************************************************************************************************************
*****************************************************************************************************************
*****************************************************************************************************************



# MCH Jiggling Trials - Decapsulated vs Control
```{r}
library(readxl)
jelly <- read_excel("~/Desktop/Publications/Manuscripts/In Prep/Comparative Egg Clutch Biomechanics/Comparative Egg Clutch Biomechanics.xlsx", 
    sheet = "MCH Jiggling Trials", na = "NA")

jelly
```

##Proportion Hatched

Summarize proportion data
```{r}
library(dplyr)
proportion_hatched = jelly %>%
  group_by(Treatment) %>%
  count(Hatched) %>% 
  mutate(prop = n/105) %>%
  filter(Hatched == "Hatched")
proportion_hatched
```

### Figure

Proportion hatched
```{r}
library(ggplot2)
library(dplyr)

proportion_hatched_summary = jelly %>%
  group_by(Treatment, Clutch) %>%
  count(Hatched) %>% 
  mutate(prop = ifelse(Hatched == "Hatched", n/3, 0)) %>%
  filter(Hatched == "Hatched" | Hatched == "Not hatched" & n == 3)
proportion_hatched_summary


proportion_hatched3=ggplot(proportion_hatched_summary, aes(y=prop, x=Treatment)) +
  lemon::geom_pointpath(aes(fill=Treatment, group=Clutch), position = position_jitter(width = .2, height = .05), size=2.5,  shorten = F, distance=0, color="black", pch=21) +
  labs(y="Proportion hatched") + #creat labels... 
  theme_classic(base_size = 9) +
  theme(legend.position="none", axis.title.y = element_text(size = 9), axis.text = element_text(color="black", size = 9), axis.ticks = element_line(color = "black"), axis.title.x = element_blank(), axis.text.x = element_blank()) +
  scale_y_continuous(breaks= c(0,.33,.67,1) ) + # allows for making "pretty" breaks on y axis... shows 5 breaks
  scale_fill_manual(values = c("darkgreen", "dimgray")) +
  geom_text(aes(y= 1.1, x = 1.5, label= "***"), color= "gray", vjust = 0, size = 4) + #add significance star/astrix 
  geom_segment(aes(x = 1, y = 1.1, xend = 2, yend = 1.1), color="gray", size = .25) # add line segment 
proportion_hatched3
```


### STATS
```{r}
prop.test(x=c(53,105), n=c(88,105))
```


2-sample test for equality of proportions with continuity correction

data:  c(53, 105) out of c(88, 105)
*X-squared = 48.368, df = 1, p-value = 3.532e-12*
alternative hypothesis: two.sided
95 percent confidence interval:
 -0.5104287 -0.2850258
sample estimates:
   prop 1    prop 2 
0.6022727 1.0000000 


t-test
```{r}
library(lme4)
library(car)
library(lmerTest)


m = lm(prop~Treatment, data = proportion_hatched_summary)

summary(m)
Anova(m)

#Check for normality assumptions
#homogeneity of variance assumption
plot(fitted(m), residuals(m))
leveneTest(prop~Treatment, data = proportion_hatched_summary)
#normality assumption
hist(residuals(m))
qqnorm(residuals(m))
qqline(residuals(m))
shapiro.test(residuals(m))


wilcox.test(prop~Treatment, data = proportion_hatched_summary)
```
Shapiro-Wilk normality test

data:  residuals(m)
W = 0.83365, p-value = 1.445e-06

Wilcoxon rank sum test with continuity correction

data:  prop by Treatment
*W = 306.5, p-value = 0.0001006*
alternative hypothesis: true location shift is not equal to 0






## Proportion Attempted to Hatch

Summarize proportion data
```{r}
library(dplyr)
proportion_attmpted_hatched = jelly %>%
  group_by(Treatment) %>%
  count(Attempted_to_Hatch) %>% 
  mutate(prop = n/105)
proportion_attmpted_hatched
```
68 embryos attempted to hatch in control treatment
88 attempted to hatch (and succedded, as above) in de-jelied treatment

### STATS
```{r}
prop.test(x=c(68,105), n=c(88,105))
```


2-sample test for equality of proportions with continuity correction

data:  c(68, 105) out of c(88, 105)
*X-squared = 24.232, df = 1, p-value = 8.54e-07*
alternative hypothesis: two.sided
95 percent confidence interval:
 -0.3252740 -0.1292715
sample estimates:
   prop 1    prop 2 
0.7727273 1.0000000 


t-test
```{r}
library(lme4)
library(car)
library(lmerTest)

proportion_attempted_to_hatch_summary = jelly %>%
  group_by(Treatment, Clutch) %>%
  count(Attempted_to_Hatch) %>% 
  mutate(prop_attemp = ifelse(Attempted_to_Hatch == 1, n/3, 0)) %>%
  filter(Attempted_to_Hatch == 1 | Attempted_to_Hatch == 0 & n == 3)
proportion_attempted_to_hatch_summary

m = lm(prop_attemp~Treatment, data = proportion_attempted_to_hatch_summary)

summary(m)
Anova(m)

#Check for normality assumptions
#homogeneity of variance assumption
plot(fitted(m), residuals(m))
leveneTest(prop_attemp~Treatment, data = proportion_attempted_to_hatch_summary)
#normality assumption
hist(residuals(m))
qqnorm(residuals(m))
qqline(residuals(m))
shapiro.test(residuals(m))


wilcox.test(prop_attemp~Treatment, data = proportion_attempted_to_hatch_summary)
```
Shapiro-Wilk normality test

data:  residuals(m)
W = 0.79943, p-value = 2.309e-08

Wilcoxon rank sum test with continuity correction

data:  prop_attemp by Treatment
*W = 377, p-value = 0.002168*
alternative hypothesis: true location shift is not equal to 0



*****************************************************************************************************************
*****************************************************************************************************************
*****************************************************************************************************************

## Hatching Complications
```{r}
proportion_hatchcomplication = jelly %>%
  group_by(Treatment) %>%
  count(Hatching_Complication)
proportion_hatchcomplication

prop.test(x=c(0,105), n=c(15,105))
```
15/105 control (14.3%)  had issues while 0 dejellied had issues



## Latency to Hatch

Latency to hatch RAW

Summarize
```{r}
library(dplyr)

jelly_summary = jelly %>% 
  group_by(Treatment, Clutch) %>%
  filter(Clutch != 49, Clutch != 54, Clutch != 57 , Clutch != 112 , Clutch != 165 , Clutch != 167 , Clutch != 169 , Clutch != 174) %>% # this is getting rid of clutches without both jellied and control values
summarise(mean_latency = mean(Latency_min, na.rm=T), SE= sd(Latency_min, na.rm=T)/sqrt(length(Latency_min)), SD = sd(Latency_min, na.rm=T), min= min(Latency_min, na.rm=T), max= max(Latency_min, na.rm=T), sample_size=n()) %>%
  filter(mean_latency != "NaN")
jelly_summary

jelly_summary2 = jelly %>% 
  group_by(Treatment) %>%
  filter(Clutch != 49, Clutch != 54, Clutch != 57 , Clutch != 112 , Clutch != 165 , Clutch != 167 , Clutch != 169 , Clutch != 174) %>% # this is getting rid of clutches without both jellied and control values
summarise(mean_latency = mean(Latency_min, na.rm=T), SE= sd(Latency_min, na.rm=T)/sqrt(length(Latency_min)), SD = sd(Latency_min, na.rm=T), min= min(Latency_min, na.rm=T), max= max(Latency_min, na.rm=T), sample_size=n()) %>%
  filter(mean_latency != "NaN")

jelly_summary2

jelly_summary3 = jelly %>% 
  group_by(Clutch) %>%
  filter(Clutch != 49, Clutch != 54, Clutch != 57 , Clutch != 112 , Clutch != 165 , Clutch != 167 , Clutch != 169 , Clutch != 174) %>% # this is getting rid of clutches without both jellied and control values
summarise(mean_latency = mean(Latency_min, na.rm=T), SE= sd(Latency_min, na.rm=T)/sqrt(length(Latency_min)), SD = sd(Latency_min, na.rm=T), min= min(Latency_min, na.rm=T), max= max(Latency_min, na.rm=T), sample_size=n()) %>%
  filter(mean_latency != "NaN")
jelly_summary3
```

### Figure

Latency Summarized Reaction Norm
```{r}
library(ggplot2)
library(scales)
mean_latency = ggplot(jelly_summary, aes(x=Treatment, y=mean_latency)) + #add the axis
  geom_errorbar(aes(ymin=mean_latency-SE, ymax=mean_latency+SE, group=Clutch), position=position_dodge(.3), width= .5, size= .25, color = "black") + 
  geom_line(aes(group=Clutch), size = 1, position = position_dodge(.3), alpha=0.4 ) +
  geom_point(aes(color=Treatment, fill= Treatment, group=Clutch), position = position_dodge(.3), size = 2.5, color = "black", pch=21) +
  labs(y="Latency to hatch (min)", x=element_blank()) + #create labels... 
  theme_classic(base_size=9) +
  theme(legend.text = element_text(size = 9), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 9), axis.ticks = element_line(color = "black")) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= "none") +
  scale_y_continuous(breaks= pretty_breaks(n=5)) + 
  scale_color_manual(values = c("darkgreen", "dimgray")) +
  scale_fill_manual(values = c("darkgreen", "dimgray")) +
  scale_color_brewer(palette = "Set2") +
  geom_text(aes(y= 6, x = 1.5, label= "***"), color= "gray", vjust = 0, size = 4) + #add significance star/astrix 
  geom_segment(aes(x = 1, y = 6, xend = 2, yend = 6), color="gray", size = .25) # add line segment 
mean_latency

#3x3 looks good in PDF
```

### STATS
```{r}
library(lme4)
library(car)
library(lmerTest)

jellypaired = jelly %>%
    filter(Clutch != 49, Clutch != 54, Clutch != 57 , Clutch != 112 , Clutch != 165 , Clutch != 167 , Clutch != 169 , Clutch != 174) # this is getting rid of clutches without both jellied and control values
jellypaired

m = lm(Latency_min~Treatment, data = jellypaired)

summary(m)
Anova(m)

#Check for normality assumptions
#homogeneity of variance assumption
plot(fitted(m), residuals(m))
leveneTest(Latency_min~Treatment, data = jellypaired)
#normality assumption
hist(residuals(m))
qqnorm(residuals(m))
qqline(residuals(m))
shapiro.test(residuals(m))


wilcox.test(Latency_min~Treatment, data = jellypaired)
```

Shapiro-Wilk normality test

data:  residuals(m)
W = 0.82758, p-value = 1.089e-10


Wilcoxon rank sum test with continuity correction

data:  Latency_min by Treatment
*W = 2452.5, p-value = 5.074e-05*
alternative hypothesis: true location shift is not equal to 0


# Figure 7
```{r}
library(gridExtra)
library(cowplot)

#  @ portrait at 7.08661 x 8.26772 max for JEB
# 4x8.5 portrait looks good

Figure6 = plot_grid(proportion_hatched3, mean_latency, labels = "AUTO", label_x = -.01, label_y = 1, scale =1, ncol=1, nrow = 2, label_size = 12, align = "v")
Figure6

# @ 3.46 x 6
```

